---
title: "Reproducible-Research-Project 1"
output: html_document
 
---
###Code for reading in the dataset and/or processing the data

```{r, echo = TRUE}
setwd("C:/Users/sravy.DESKTOP-B0LIHLK/Desktop/Datascience/RR Week2")
activity<-read.csv("activity.csv")
```

```{r} 
head(activity)
str(activity)
```

###Histogram of the total number of steps taken each day

###Ignoring NA Values
```{r}
data <- activity[!(is.na(activity$steps)), ]
```
###Total steps taken per day
```{r}
totalSteps <- aggregate(steps ~ date, data, sum)
head(totalSteps)
```
###Create Histogram
```{r}
paletteBlue <- colorRampPalette(c("skyblue", "darkblue", "skyblue"))
hist(totalSteps$steps, breaks=20, xlab="Number of Steps Taken", 
     main="Histogram of the Total Number of Steps Taken per Day",
     col=paletteBlue(22), family="serif")
```

###Mean and median number of steps taken each day


```{r, message=FALSE}
library(dplyr)
totalStepsSummary <- summarise(totalSteps, meanOfTotalSteps=mean(totalSteps$steps),
                               medianOfTotalSteps=median(totalSteps$steps))
print(totalStepsSummary)
```

###Time series plot of the average number of steps taken

```{r}
meanStepsInterval <- aggregate(steps ~ interval, data, mean)
head(meanStepsInterval)
```
### Plotting a timeseries of the mean steps per interval
```{r}
plot(x=meanStepsInterval$interval, y=meanStepsInterval$steps, type="l",
     main="Time Series Plot of Average Steps Taken per Interval",
     ylab="Number of Steps", xlab="Intervals (in 5 mins)",
     col="darkblue", lwd=1.5, family="serif")
```

###The 5-minute interval that, on average, contains the maximum number of steps
```{r}
meanStepsInterval[grep(max(meanStepsInterval$steps), meanStepsInterval$steps), ]
```

###Code to describe and show a strategy for imputing missing data

```{r}
anyNA(activity)
```

###Analyze NA Values
```{r}
data.frame(steps=sum(is.na(activity$steps)), 
           interval=sum(is.na(activity$interval)), 
           date=sum(is.na(activity$date)))
```
###2304 missing values in Steps

###Using Mean/mode value substitutions

```{r}

imputedData <- activity
for(x in 1:17568) {
    if(is.na(imputedData[x, 1])==TRUE) {
        imputedData[x, 1] <- meanStepsInterval[meanStepsInterval$interval %in% imputedData[x, 3], 2]
    }
}
head(imputedData)
```
###Check for any NA values
```{r}
anyNA(imputedData)
```

###Histogram of the total number of steps taken each day after missing values are imputed
```{r}
imputedTotalStepsDay <- aggregate(steps ~ date, imputedData, sum)
head(imputedTotalStepsDay)
```
###Plot Histogram

```{r}
paletteRed <- colorRampPalette(c("deeppink", "darkred", "deeppink"))
hist(imputedTotalStepsDay$steps, breaks=20, xlab="Number of Steps Taken", 
     main="Histogram of Total Number of Steps Taken per Day (With Imputed Values)",
     col=paletteRed(22), family="serif")
```

###Calculate and report the mean and median total number of steps taken per day with imputed data

```{r}
imputedStepsSummary <- summarise(imputedTotalStepsDay, 
                                 meanOfTotalSteps=mean(imputedTotalStepsDay$steps), 
                                 medianOfTotalSteps=median(imputedTotalStepsDay$steps))  
print(imputedStepsSummary)
```
###compare the steps with and without imputed data
```{r}
rbind(totalStepsSummary, imputedStepsSummary)
```
### the values look very similar if not exact due to the mean/mode value imputing strategy

###Histogram comparision of imputes and non-imputed data sets
```{r}
par(mfrow = c(1, 2))
hist(totalSteps$steps, breaks=20, xlab="Number of Steps Taken", 
     col=paletteBlue(22), family="serif", ylim=c(0, 20), main=NULL)
hist(imputedTotalStepsDay$steps, breaks=20, xlab="Number of Steps Taken", 
     col=paletteRed(22), family="serif", ylim=c(0, 20), main=NULL)
mtext("Histograms of Total Number of Steps Taken per Day, Without/With Imputed Values",
      adj=0.95, family="serif", font=2)
```

###Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

###Weekday function to calculate the day of the week

```{r}
weekday.or.weekend <- function(date) {
    day <- weekdays(date)
    if (day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))
        return("weekday")
    else if (day %in% c("Saturday", "Sunday"))
        return("weekend")
    else
        stop("invalid date")
}
imputedData$date <- as.Date(imputedData$date)
imputedData$day <- sapply(imputedData$date, FUN=weekday.or.weekend)
```

###Panel Plot
```{r}
library(ggplot2)
averages <- aggregate(steps ~ interval + day, data=imputedData, mean)
ggplot(averages, aes(interval, steps)) + geom_line() + facet_grid(day ~ .) +
    xlab("5-minute interval") + ylab("Number of steps")
```